---
title: "HW3 - Lucas Fellmeth, Sven Bergmann"
date: "2023-10-18"
output: pdf_document
latex_engine: xelatex
---

```{r setup, echo = F}
library(knitr)
opts_chunk$set(tidy.opts = list(width.cutoff = 80), tidy = TRUE)
```

# Chapter 8

## Problem 4

The points-per-game statistics from the 1993 NBA season were analyzed for basketball players who went to college in four particular ACC schools:
Duke, North Carolina, North Carolina State, and Georgia Tech.
We want to find out if scoring is different for the players from different schools.
Can this be analyzed with a parametric procedure?
Why or why not?
The classical $F$-test that assumes normality of the populations yields $F = 0.41$ and $H_0$ is not rejected.
What about the nonparametric procedure?

| Duke | UNC  | NSCU | GT   |
|------|------|------|------|
| 7.5  | 5.5  | 16.9 | 7.9  |
| 8.7  | 6.2  | 4.5  | 7.8  |
| 7.1  | 13.0 | 10.5 | 14.5 |
| 18.2 | 9.7  | 4.4  | 6.1  |
|      | 12.9 | 4.6  | 4.0  |
|      | 5.9  | 18.7 | 14.0 |
|      | 1.9  | 8.7  |      |
|      |      | 15.8 |      |

### Solution

```{r}
duke <- c(7.5, 8.7, 7.1, 18.2)
unc <- c(5.5, 6.2, 13.0, 9.7, 12.9, 5.9, 1.9)
nscu <- c(16.9, 4.5, 10.5, 4.4, 4.6, 18.7, 8.7, 15.8)
gt <- c(7.9, 7.8, 14.5, 6.1, 4.0, 14.0)

data <- c(duke, unc, nscu, gt)
belong <- c(rep(1, length(duke)), rep(2, length(unc)), rep(3, length(nscu)), rep(4, length(gt)))

kruskal.test(data, belong)
```

## Problem 5

Some varieties of nematodes (roundworms that live in the soil and are frequently so small they are invisible to the naked eye) feed on the roots of lawn grasses and crops such as strawberries and tomatoes.
This pest, which is particularly troublesome in warm climates, can be treated by the application of nematocides.
However, because of size of the worms, it is difficult to measure the effectiveness of these pesticides directly.
To compare four nematocides, the yields of equal-size plots of one variety of tomatoes were collected.
The data (yields in pounds per plot) are shown in the table.
Use a nonparametric test to find out which nematocides are different:


| Nematocide A | Nematocide B | Nematocide C | Nematocide D |
|--------------|--------------|--------------|--------------|
| 18.6         | 18.7         | 19.4         | 19.0         |
| 18.4         | 19.0         | 18.9         | 18.8         |
| 18.4         | 18.9         | 19.5         | 18.6         |
| 18.5         | 18.5         | 19.1         | 18.7         |
| 17.9         |              | 18.5         |              |

### Solution


# Chapter 9

## Problem 15

West of Tokyo lies a large alluvial plain, dotted by a network of farming villages.
Matui (1968) analyzed the position of the 911 houses making up one of those villages.
The area studied was a rectangle, $3 km \times 4 km$.
A grid was superimposed over a map of the village, dividing its 12 $km^2$ into 1200 plots, each 100 m on a side.
The number of houses on each of those plots was recorded in a $30 \times 40$ matrix of data.
Test the hypothesis that the distribution of number of houses per plot is Poisson.
Use $\alpha = 0.05$:

| Number    | 0   | 1   | 2   | 3  | 4 | $\geq$ 5 |
|-----------|-----|-----|-----|----|---|----------|
| Frequency | 584 | 398 | 168 | 35 | 9 | 6        |

*Hint:*
Assume that parameter $\lambda = 0.76$ (approximately the ratio 911/1200).
Find theoretical frequencies first.
For example, the theoretical frequency for number = 2 is $np_2 = 1200 \times 0.762^2/2! \times \exp{(-.76)} = 162.0745$, while the observed frequency is 168.
Subtract an additional degree of freedom because $\lambda$ is estimated from the data (Figure 9.6).

### Solution