---
title: "HW4 - Lucas Fellmeth, Sven Bergmann"
date: "2023-11-01"
output: pdf_document
latex_engine: xelatex
---

```{r setup, echo = F}
library(knitr)
opts_chunk$set(tidy.opts = list(width.cutoff = 80), tidy = TRUE)
```

# Chapter 11

# Exercise 5
Recall Exercise 6.3 based on 100 measurements of the speed of light in air.
In that chapter, we tested the data for normality.
Use the same data to construct a density estimator that you feel gives the best visual display of the information provided by the data.
What parameters did you choose?
The data can be downloaded from
http://www.itl.nist.gov/div898/strd/univ/data/Michelso.dat


```{r}
x <- as.numeric(read.delim2("Michelso.dat.txt")[[1]])
hist(x, freq = F, nclass = 10)
kde <- density(x, bw = "SJ")
lines(kde$x, kde$y, col = "red", lwd = 2)
```

# Chapter 12

# Exercise 1
Using robust regression, find the intercept and slope $\tilde{\beta}_0$ and $\tilde{\beta}_1$ for each of the four data sets of Anscombe (1973) from p. 244.
Plot the ordinary least-squares regression along with the rank regression estimator of slope.
Contrast these with one of the other robust regression techniques.
For which set does $\tilde{\beta}_1$ differ the most from its LS counterpart $\hat{\beta}_1 = 0.5$?
Note that in the fourth set, 10 out of 11 Xs are equal, so one should use
$$S_{ij} = (Y_j - Y_i)/(X_j - X_i + \epsilon)$$ to avoid dividing by 0.
After finding $\tilde{\beta}_0$ and $\tilde{\beta}_1$, are they different than $\hat{\beta}_0$ and $\hat{\beta}_1$?
Is the hypothesis $H_0 : \beta_1 = 1/2$ rejected in a robust test against the alternative $H_1 : \beta_1 < 1/2$, for data set 3?
Note here $\beta_{10} = 1/2$.




Exercises 11.5 and 12.1 (here, for each data set, compare the least squares estimator with the least median of squares estimator and the M-estimator with Tukey's bisquare function). Due Wednesday Nov 8.