---
title: "HW2 - Lucas Fellmeth, Sven Bergmann"
date: "2023-10-05"
output: pdf_document
latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Chapter 7

## Problem 3

Diet A was given to a group of 10 overweight boys between the ages of 8 and 10.
Diet B was given to another independent group of 8 similar overweight boys.
The weight loss is given in the table below.
Using WMW test, test the hypothesis that the diets are of comparable effectiveness against the two-sided alternative.
Use $\alpha = 5\%$ and normal approximation.

|        |   |   |   |    |   |   |   |   |   |   |
|--------|---|---|---|----|---|---|---|---|---|---|
| Diet A | 7 | 2 | 3 | -1 | 4 | 6 | 0 | 1 | 4 | 6 |
| Diet B | 5 | 6 | 4 | 7  | 8 | 9 | 7 | 2 |   |   |

### Solution

We test for $H_0 : F_A(x) = F_B(x)$.

First we find
$$W_n =\sum_{n=1}^{n_1} R_i$$

We reject when $W_n \geq k_{0.05}$.

```{r}
#rm(list=ls())
a <- c(7, 2, 3, -1, 4, 6, 0, 1, 4, 6)
b <- c(5, 6, 4, 7, 8, 9, 7, 2)

n_1 <- length(a)
n_2 <- length(b)
c <- c(a, b)
n <- length(c)
rnk <- rank(c)
W_n <- sum(rnk[1:n_1]) # sum of the ranks of a
U <- W_n - (n_1 * (n_1 + 1) / 2)
```

Since there are ties in the ranks
$$W_n \sim \mathcal{N} \left(
\frac{n_1(n+1)}{2},
\frac{n_1n_2(n+1)}{12} - \frac{n_1n_2\sum_{i=1}^{k}(t_i^3 -t_i)}{12n(n+1)}
\right)$$
where $t_1, \ldots,t_k$ are the number of different observations among all the observations in the combined sample.


We reject if
\begin{align*}
\frac{W_n -\frac{n_1(n+1)}{2}}{\sqrt{\frac{n_1n_2(n+1)}{12} - \frac{n_1n_2\sum_{i=1}^{k}(t_i^3 -t_i)}{12n(n+1)}}} &> z_{0.05} \\
\Leftrightarrow W_n  &> z_{0.05} +\frac{n_1(n+1)}{2}* \left( \sqrt{\frac{n_1n_2(n+1)}{12} - \frac{n_1n_2\sum_{i=1}^{k}(t_i^3 -t_i)}{12n(n+1)}} \right)
\end{align*}

```{r}
t <- tabulate(rnk)
t <- t[t != 0]
t_sum <- sum(t^3 - t)

lower <- (1 - pnorm(0.05)) *
  (n_1 * (n + 1) / 2) *
  sqrt(n_1 * n_2 * (n + 1) / 12 - n_1 * n_2 * t_sum / (12 * n * (n + 1)))

if (W_n > lower) {
  print("We reject the null hypothesis fot he alternative")
} else {
  print("We accept the null hypothesis")
}
```

## Problem 6

Use the link below to see the results of an experiment on the effect of prior information
on the time to fuse random dot stereograms.
One group (NV) was given either no information or just verbal information about the shape of the embedded object.
A second group (VV) received both verbal information and visual information (e.g. a drawing of the object).
Does the median time prove to be greater for the NV group?
Compare your results to those from a two-sample t-test.

### Solution

This time, our null hypothesis is:
$$H_0:\mu_{NV}^2 = \mu_{VV}^2 \text{ vs } H_1:\mu_{NV}^2 >\mu_{VV}^2$$

```{r}
#rm(list = ls())
stereogram_data <- read.delim2("stereograms.txt")
stereogram_data$fusion.time <- as.numeric(stereogram_data$fusion.time)
stereogram_data$group <- factor(stereogram_data$group)
table(stereogram_data$group) # we have n_1 and n_2 > 10
#means <- with(stereogram_data, tapply(fusion.time, group, mean))
means <- aggregate(fusion.time ~ group, stereogram_data, mean)
print(means)

```
Yes, the median time does prove to be greater for the NV group.

```{r}

t.test(subset(stereogram_data, subset = (group =="NV"))$fusion.time , subset(stereogram_data, subset = (group =="VV"))$fusion.time, alternative = "greater", var.equal = FALSE)

```

The two-sample t-test confirms our observation. We rejcect $H_0$ for the alternative.

# Problem 7

Derive the exact distribution of the Mann–Whitney U statistic in the case that $n_1 = 4$ and $n_2 = 2$.

## Solution

The exact distribution can be found as
$$P(W_n = m) = \frac{k_{n_1,n_2}(m)}{\left( \begin{array}{c} n \\ n_1 \end{array} \right)}, \frac{n_1(n_1+1)}{2} \leq m \leq \frac{n_1(2n -n_1 +1)}{2}$$

with $k_{n_1,n_2} (m)$ the number of all arrangements of zeroes and ones in $(S_1(X,Y), \ldots, S_n(X,Y))$
such that $W_n =\sum_{i=1}^n iS_i(X,Y) = m$.

We have
$$n_1 =4, n_2= 2, n = n_1 +n_2 = 6$$
so there are $\left( \begin{array}{c} 6 \\ 4\end{array} \right) = \left( \begin{array}{c} 6 \\ 2 \end{array} \right) = 15$ distinguishable configurations of the vector $(S_1, S_2,\ldots, S_6)$.


The minimum of $W_6$ is $10$ and the maximum is $18$.


Then we get the following distribution:


\begin{table}[ht]
\centering
\caption{Distribution of $W_6$ with $n_1=4$ and $n_2=2$}
\begin{tabular}[t]{llc}
\hline
$W_6$ & Configuration &Probability\\
\hline
$10$ &$(1,2,3,4)$ & 1/15\\
$11$ &$(1,2,3,5)$ & 1/15\\
$12$ &$(1,2,3,6), (1,2,4,5)$ & 2/15\\
$13$ &$(1,2,4,6), (1,3,4,5)$ & 2/15\\
$14$ &$(1,3,4,6), (2,3,4,5), (1,2,5,6)$ & 3/15\\
$15$ &$(2,3,4,6), (1,3,5,6)$ & 2/15\\
$16$ &$(2,3,5,6), (1,4,5,6)$ & 2/15\\
$17$ &$(2,4,5,6)$ & 1/15\\
$18$ &$(3,4,5,6)$ & 1/15\\
\hline
\end{tabular}
\end{table}%

