---
title: "HW6 - Lucas Fellmeth, Sven Bergmann"
date: "2023-11-29"
output: pdf_document
latex_engine: xelatex
---

```{r setup, echo = F}
library(knitr)
opts_chunk$set(tidy.opts = list(width.cutoff = 80), tidy = TRUE)
```
# 1
The goal of this problem is to estimate the regression function of acceleration vs time for the `mcycle` data in the package `MASS`, using cubic smoothing splines.

```{r}
library(MASS)
library(splines)
```

## A
For a reasonable range of smoothing parameters $\lambda$, compute and plot the generalized cross validation measure $GCV(\lambda)$ and find the optimal smoothing parameter.

```{r, out.width = "50%", out.height = "50%"}
with(mcycle, {
  lambda <- seq(10^(-6), 10^(-3), by = 10^(-6))

  smoothing <- function(l, x, y) {
    return(smooth.spline(x, y, lambda = l)$cv.crit)
  }

  gcv <- sapply(lambda, smoothing, x = accel, y = times)
  tmp <- data.frame(lambda = lambda, gcv = gcv)
  lambda_opt <<- tmp[which.min(tmp$gcv),]$lambda
  plot(tmp, type = 'l')
  plot.new()
  plot(times, accel, main = paste('min_lambda =', lambda_opt))
  lines(smooth.spline(x = times, y = accel, lambda = lambda_opt), col = 'red', type = 'l')
})
```

## B
Using the Demmler-Reinsch basis representation, how many basis functions contribute significantly to the fit, as determined, for instance, by accumulating $90\%$ of the total degrees of freedom?

https://statisticaloddsandends.wordpress.com/2019/10/04/the-demmler-reinsch-basis-for-smoothing-splines/
```{r}
library(eBsc)
with(mcycle, {
  degrees_of_freedom <- smooth.spline(accel, times, lambda = lambda_opt)$df
  basis <- drbasis(nn = length(accel), qq = 3)
  pca <- prcomp(basis$eigenvectors)
  # summary(pca)
  with(basis, {
    times <- as.matrix(times)
    print(dim(eigenvalues))
    print(dim(t(eigenvectors) %*% times))
    print(dim(t(eigenvectors)))
    print(t(eigenvalues) %*% (eigenvectors %*% times))
    # sum(t(eigenvalues) %*%
    #       (eigenvectors %*% times) *
    #       eigenvectors)
    # print(eigenvectors)
  })
})
```
