---
title: "HW5 - Lucas Fellmeth, Sven Bergmann"
date: "2023-11-16"
output: pdf_document
latex_engine: xelatex
---

```{r setup, echo = F}
library(knitr)
opts_chunk$set(tidy.opts = list(width.cutoff = 80), tidy = TRUE)
```
## 1
The goal of this problem is to estimate the regression function of acceleration vs time for the `mcycle` data in the package `MASS`.

### A
Show that the Nadaraya-Watson estimator can be expressed as $\hat{Y} = HY$. Find the "hat matrix" $H$ explicitly.

We know that the Nadaraya-Watson estimator of $\hat{m}(x_i)$ is defined by
$$\hat{m}(x_i) = \frac{\sum_{j=1}^n K_h(x_j - x_i)Y_j}{\sum_{k=1}^n K_h(x_k - x_i)},$$
where
$$K_h(x) = \frac{1}{h}K\left(\frac{x}{h}\right)$$
with $h$ as associated bandwidth.

So
\begin{align*}
\hat{m}(x_i)  &= \frac{\sum_{j=1}^n K_h(x_j - x_i)Y_j}{\sum_{k=1}^n K_h(x_k - x_i)} \\
&= \frac{\sum_{j=1}^n \frac{1}{h} K(\frac{x_j - x_i}{h})Y_j}{\sum_{k=1}^n \frac{1}{h} K(\frac{x_k - x_i}{h})} \\
&= \frac{\frac{1}{h} \sum_{j=1}^n K(\frac{x_j - x_i}{h})Y_j}{\frac{1}{h} \sum_{k=1}^n K(\frac{x_k - x_i}{h})} \\
&= \frac{\sum_{j=1}^n K(\frac{x_j - x_i}{h})Y_j}{\sum_{k=1}^n K(\frac{x_k - x_i}{h})} \\
&= \frac{\sum_{j=1}^n K(\frac{x_j - x_i}{h})}{\sum_{k=1}^n K(\frac{x_k - x_i}{h})} Y_j \\
&= \sum_{j=1}^n \underbrace{\frac{K(\frac{x_j - x_i}{h})}{\sum_{k=1}^n K(\frac{x_k - x_i}{h})}}_{H_{ij}} Y_j \\
\implies
\underbrace{\left(\begin{array}{c}
\hat{m}(x_1) \\
\vdots \\
\hat{m}(x_n)
\end{array}\right)}_{\hat{Y}}
&= \underbrace{\left(\begin{array}{ccc}
H(x_{11}) & \ldots & H(x_{1n}) \\
\vdots & \ddots & \vdots \\
H(x_{n1}) & \ldots & H(x_{nn})
\end{array}\right)}_{H}
\cdot
\underbrace{\left(\begin{array}{c}
m(x_1) \\
\vdots \\
m(x_n)
\end{array}\right)}_{Y}
\end{align*}


### B
For a reasonable range of bandwidths $h$, compute and plot the generalized cross validation measure $GCV(h)$ and find the optimal bandwidth.

```{r}
library(MASS)
library(splines)
```
First, we implement the generalized cross validation measure $GCV(h)$ which is defined by
$$GCV(h) = \frac{1}{n} \sum_{i=1}^n \left[\frac{Y_i - \hat{m}_h(x)}{1 - \frac{tr H(h)}{n}}\right]^2$$
where
```{r GCV(h)}
GCV <- function(x, y, h, kernel = "normal") {
  y_hat <- ksmooth(x, y, kernel, bandwidth = h)$y
  H <- y_hat %*% t(y)
  H_trace <- sum(diag(H))
  n <- length(y)
  gcv <- mean((y - y_hat)^2) / (1 - (H_trace / n))^2
  return(gcv)
}
```
```{r}
with(mcycle, {
  bw <- 1:15
  gcv <- sapply(bw, GCV, x = times, y = accel)
  plot(bw, gcv, type = 'b')
  plot.new()
  plot(times, accel)
  lines(times, ksmooth(times, accel, kernel = 'normal', bandwidth = min(gcv))$y, col = 'red', type = 'l')
})
```
