---
title: "HW5 - Lucas Fellmeth, Sven Bergmann"
date: "2023-11-16"
output: pdf_document
latex_engine: xelatex
---

```{r setup, echo = F}
library(knitr)
opts_chunk$set(tidy.opts = list(width.cutoff = 80), tidy = TRUE)
```
## 1
The goal of this problem is to estimate the regression function of acceleration vs time for the `mcycle` data in the package `MASS`.

### A
Show that the Nadaraya-Watson estimator can be expressed as $\hat{Y} = HY$. Find the "hat matrix" $H$ explicitly.

We know that the Nadaraya-Watson estimator of $\hat{m}(x)$ is defined by
$$\hat{m}(x) = \frac{\sum_{i=1}^n K_h(X_i - x)Y_i}{\sum_{i=1}^n K_h(X_i - x)},$$
where
$$K_h(x) = \frac{1}{h}K\left(\frac{x}{h}\right)$$
with $h$ as associated bandwidth.



### B
For a reasonable range of bandwidths $h$, compute and plot the generalized cross validation measure $GCV(h)$ and find the optimal bandwidth.

```{r}
library(MASS)
library(splines)
```
First, we implement the generalized cross validation measure $GCV(h)$ which is defined by
$$GCV(h) = \frac{1}{n} \sum_{i=1}^n \left[\frac{Y_i - \hat{m}_h(x)}{1 - \frac{tr S(h)}{n}}\right]^2$$
where
```{r GCV(h)}
GCV <- function(h) {

}
```
```{r}
with(mcycle, {
  plot(times, accel)
  for (b in 1:10) {
    lines(ksmooth(times, accel, "normal", bandwidth = b), col = "red")
  }
})
```
